---
title: "Using apihelperR"
author: "Venkata Balaji Anupoju and Marwan Karim"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Using apihelperR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
body {
  max-width: 900px;
  margin: 0 auto;
  padding: 25px 35px;
  line-height: 1.6;
}

h1, h2 {
  margin-top: 2em;
}

pre {
  border-radius: 6px;
  padding: 12px;
}

img {
  display: block;
  margin: 25px auto 40px auto;
  max-width: 72%;
  height: auto;
}
</style>

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  out.width = "72%"
)
```

# Introduction

The `apihelperR` package provides functions for retrieving earthquake
data from the USGS Earthquake API.

The main function is `get_earthquakes()`. It allows the user to specify
a date range, minimum earthquake magnitude, and maximum number of
results.

The function returns the earthquake information as a data frame that
can be used for further analysis in R.

# Load the package

```{r}
library(apihelperR)
```

# Get earthquake data

Earthquake data can be retrieved using `get_earthquakes()`.

```{r, eval=FALSE}
quakes <- get_earthquakes(
  start_date = "2026-01-01",
  end_date = "2026-01-02",
  min_magnitude = 4,
  limit = 10
)

quakes
```

The `start_date` and `end_date` arguments define the time period to
search.

The `min_magnitude` argument specifies the minimum earthquake magnitude,
while `limit` controls the maximum number of earthquakes returned.

# Inspect the returned data

The first rows of the returned data can be inspected using `head()`.

```{r, eval=FALSE}
head(quakes)
```

The structure of the data can also be examined using `str()`.

```{r, eval=FALSE}
str(quakes)
```

The returned data frame contains the following variables:

- `id`
- `time`
- `magnitude`
- `place`
- `longitude`
- `latitude`
- `depth`
- `tsunami`

These variables provide information about when and where each
earthquake occurred together with its magnitude, depth, and tsunami
status.

# Input validation

The function checks the supplied arguments before sending a request to
the API.

For example, the start date cannot be later than the end date.

```{r, eval=FALSE}
get_earthquakes(
  start_date = "2026-01-10",
  end_date = "2026-01-01"
)
```

This produces an error because the date range is invalid.

The `limit` argument must also be between 1 and 20000.

```{r, eval=FALSE}
get_earthquakes(
  start_date = "2026-01-01",
  end_date = "2026-01-02",
  limit = 0
)
```

# Filtering earthquakes by magnitude

The `min_magnitude` argument can be used to retrieve only earthquakes
above a selected magnitude.

For example, the following query requests earthquakes with a magnitude
of at least 5.

```{r, eval=FALSE}
large_quakes <- get_earthquakes(
  start_date = "2026-01-01",
  end_date = "2026-01-07",
  min_magnitude = 5,
  limit = 20
)

head(large_quakes)
```

# Complete workflow

A complete example using the package can be performed as follows:

```{r, eval=FALSE}
quakes <- get_earthquakes(
  start_date = "2026-01-01",
  end_date = "2026-01-02",
  min_magnitude = 4,
  limit = 10
)

head(quakes)
str(quakes)
```

The `apihelperR` package therefore provides a simple interface for
retrieving earthquake information from the USGS Earthquake API and
returning it in a format suitable for analysis in R.
